quollr: An R Package for Visualizing 2D Models from Nonlinear Dimension Reductions in High Dimensional Space

### 2024-04-01 paper-quollr.Rmd
Add paper sections to intro and add bechnark argument to vis

@@

@@

@@

<!--2D model vis: triangular mesh, colored trimesh, removed edge trimesh-->
To visualize the 2D model, mainly three functions are used. As shown in Figure \@ref(fig:mesh-plots)a, `geom_trimesh()` to visualize the triangular mesh by adding a new layer to `ggplot()`. After identifying benchmark value to remove long edge, `vis_lg_mesh()` is used to visualize the triangular mesh by coloring the small and long edges. As shown in Figure \@ref(fig:mesh-plots)b, the small and long edges are colored by black and red respectively. Following this, `vis_rmlg_mesh()` is used to visualize smoothed 2D model which is the 2D model after removing the long edges (see Figure \@ref(fig:mesh-plots)c). In `vis_lg_mesh()` and `vis_rmlg_mesh()`, `benchmark_value` argument controls the edge removal in 2D.

@@

>>>>

The paper is organized as follows. Section **Methodology** details the main steps of the algorithm. In next section, introduces the implementation of `quollr` package on CRAN, including demonstration of the package's key functions and visualization capabilities. We illustrate the algorithm's functionality to study clustering data set in **Application** section, and describe a visual heuristic to describe parameter selection. Finally, we give a brief conclusion of the paper and discuss potential opportunities for use of our algorithm.

@@

trimesh_removed_s_curve_umap_0.3 <- vis_rmlg_mesh(distance_edges = distance_s_curve, benchmark_value = 0.3, tr_coord_df = tr_from_to_df_s_curve, distance_col = "distance")
trimesh_removed_s_curve_umap_0.3 <- trimesh_removed_s_curve_umap_0.3 +
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

trimesh_removed_s_curve_umap_0.8 <- vis_rmlg_mesh(distance_edges = distance_s_curve, benchmark_value = 0.8, tr_coord_df = tr_from_to_df_s_curve, distance_col = "distance")
trimesh_removed_s_curve_umap_0.8 <- trimesh_removed_s_curve_umap_0.8 +
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

@@

<!--discuss about default settings-->

@@

<!--2D model vis: triangular mesh, colored trimesh, removed edge trimesh-->
To visualize the 2D model, mainly three functions are used. As shown in Figure \@ref(fig:mesh-plots)a, `geom_trimesh()` to visualize the triangular mesh by adding a new layer to `ggplot()`. After identifying benchmark value to remove long edge, `vis_lg_mesh()` is used to visualize the triangular mesh by coloring the small and long edges. As shown in Figure \@ref(fig:mesh-plots)b, the small and long edges are colored by black and red respectively. Following this, `vis_rmlg_mesh()` is used to visualize smoothed 2D model which is the 2D model after removing the long edges (see Figure \@ref(fig:mesh-plots)c). In `vis_lg_mesh()` and `vis_rmlg_mesh()`, `benchmark_value` argument controls the edge removal in 2D. Using small value of `benchmark_value`, will produce a triangular mesh with missing data structure; for example `benchmark_value = 0.3` shows two clusters rather than continuous structure, while `benchmark_value = 0.8` creates long edges and mislead the data structure in $p-D$ space.

@@

```{r mesh-diffplots, out.width="100%", fig.cap="Visualization of model generated for UMAP embedding for S-curve data in 2D space with different benchmark value applied to remove long edges: (a) benchmark_value = 0.3, and (b) benchmark_value = 0.8. Small benchmark values will generate misleading data structures in 2D space, while; large bechmark value will generate more long edges in $p-D$ and miinterpret the $p-D$ data structure."}

trimesh_removed_s_curve_umap_0.3 + trimesh_removed_s_curve_umap_0.8 +
  plot_layout(ncol = 2) &
  theme(plot.tag = element_text(size = 8))
```
